Effect of quality-adjusted production function inputs

### 2015-05-15 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
adj-->adjusted, unadj-->unadjusted

<<setup, echo=FALSE, message=FALSE, eval=TRUE>>=
require(EconData)
require(EconModels)
require(dplyr)
require(reshape2)
require(ggplot2)
require(mosaic)

# Tell whether to re-run the analysis to create the models.
# This file will use previously-saved models to generate
# the paper if set to FALSE, greatly reducing time.
# If something changes in the data or in the code that
# performs the analysis, set 
# rerunModels = TRUE
# once to re-generate the information needed to 
# create the paper.
#####################################################
rerunModels=FALSE
#####################################################
@

@@

historicalPlot(histDataForGraph, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(1,2,4,3,2,4,3,2,3,3)) + 
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  coord_cartesian(ylim=c(0.5, 13)) +
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

# Fix levels
pre_econometric_data$flavor <- relevelFactor(as.factor(as.character(pre_econometric_data$flavor)), 
                                               c("Unadjusted", "Quality-adjusted"))
pre_econometric_data$Country <- relevelFactor(as.factor(pre_econometric_data$Country), 
                                              c("PT", "UK"))
pre_econometric_data$variable <- relevelFactor(as.factor(pre_econometric_data$variable),
                                               unique(pre_econometric_data$variable))
# Now make the plot
historicalPlot(pre_econometric_data, 
               mapping=aes(x=Year, y=ratio, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(2,4,3, 2,4,3, 2,3,3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with $(le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest == "lek" | nest=="kl")))
@

@@

<<Portugal C-D undadjusted without energy>>=
naturalCoef(models$PT$unadj$noE$CD)
@

<<Portugal C-D undadjusted with energy>>=
naturalCoef(models$PT$unadj$withE$CD)
@

The Solow residual (\texttt{lambda}) is always negative. 
However, the absolute value of \texttt{lambda} decreases from 
\Sexpr{naturalCoef(models$PT$unadj$noE$CD)$lambda} to 
\Sexpr{naturalCoef(models$PT$unadj$withE$CD)$lambda}
when adding energy.
So it appears that for Portugal, 
using unadjusted inputs and including energy as a factor of production
reduces the absolute value of the Solow residual somewhat.

@@

<<Portugal C-D adjusted without energy>>=
naturalCoef(models$PT$adj$noE$CD)
@

<<Portugal C-D adjusted with energy>>=
naturalCoef(models$PT$adj$withE$CD)
@

Using quality-adjusted inputs, the Solow residual (\texttt{lambda}) moves from 
\Sexpr{naturalCoef(models$PT$adj$noE$CD)$lambda} to
\Sexpr{naturalCoef(models$PT$adj$withE$CD)$lambda} when including energy.
The absolute value of the Solow residual again decreases when adding energy.
But, with quality-adjusted inputs, the effect is much stronger.

@@

\begin{table} \caption{Portugal Solow residual results with Cobb-Douglas.} 
\label{tab:lambda_CD_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CD)$lambda}    & \Sexpr{naturalCoef(models$PT$adj$noE$CD)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CD)$lambda}  & \Sexpr{naturalCoef(models$PT$adj$withE$CD)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{UK Solow residual results with Cobb-Douglas.} 
\label{tab:lambda_CD_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CD)$lambda}   
                                                & \Sexpr{naturalCoef(models$UK$adj$noE$CD)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CD)$lambda} 
                                                & \Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the UK, the quality-adjusted, with energy Solow residual is 
\Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda * 100}\%,
a very small number indeed.
For the UK, it appears that \emph{both} quality-adjustment and addition of energy as a factor of production
are required to effectively eliminate the Solow residual,
reducing it to \Sexpr{naturalCoef(models$UK$adj$withE$CD)$lambda}.

Adding energy alone actually \emph{in}creases the Solow residual from 
\Sexpr{naturalCoef(models$UK$unadj$noE$CD)$lambda} to
\Sexpr{naturalCoef(models$UK$unadj$withE$CD)$lambda}.
The details for each model with unadjusted factors of production are shown below.

<<UK CD Unadjusted Coefficients>>=
naturalCoef(models$UK$unadj$noE$CD)
naturalCoef(models$UK$unadj$withE$CD)
@

When energy is included, much of the effect of capital (\texttt{alpha\_1}) is 
attributed to energy (\texttt{alpha\_3}), 
as \texttt{alpha\_1} moves from 
\Sexpr{naturalCoef(models$UK$unadj$noE$CD)$alpha_1} to
\Sexpr{naturalCoef(models$UK$unadj$withE$CD)$alpha_1}.

@@

<<UK CD Quality-adjusted Coefficients>>=
naturalCoef(models$UK$adj$noE$CD)
naturalCoef(models$UK$adj$withE$CD)
@

\texttt{alpha\_1} moves from 
\Sexpr{naturalCoef(models$UK$adj$noE$CD)$alpha_1} to
\Sexpr{naturalCoef(models$UK$adj$withE$CD)$alpha_1}.

@@

<<PT CES-kl, warning=FALSE>>=
naturalCoef(models$PT$unadj$noE$CES$kl)
naturalCoef(models$PT$adj$noE$CES$kl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals,
\Sexpr{naturalCoef(models$PT$unadj$noE$CES$kl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the Solow residual.

@@

<<PT CES-kle unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CES$kle)
@

@@

<<PT CES-kle quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CES$kle)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CES$kl)$lambda} 
                                                            & \Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CES$kle)$lambda} 
                                                            & \Sexpr{naturalCoef(models$PT$adj$withE$CES$kle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CES$kl)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CES$kle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adj$withE$CES$kle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<PT CES-lek unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CES$lek)
@

@@

<<PT CES-lek quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CES$lek)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CES$kl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CES$lek)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adj$withE$CES$lek)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($le$)$k$ nesting.} 
\label{tab:sse_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CES$kl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CES$lek)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adj$withE$CES$lek)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Similar to the CES model with ($kl$)$e$ nesting,
the CES model with ($le$)$k$ nesting
produces positive Solow residuals in all cases. 
After energy has been added, quality-adjustment 
has very little effect on the Solow residual,
creating a slight increase from 
\Sexpr{naturalCoef(models$PT$unadj$withE$CES$lek)$lambda} to
\Sexpr{naturalCoef(models$PT$adj$withE$CES$lek)$lambda}.

@@

<<PT CES-ekl unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadj$withE$CES$ekl)
@

@@

<<PT CES-ekl quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adj$withE$CES$ekl)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadj$noE$CES$kl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadj$withE$CES$ekl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adj$withE$CES$ekl)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($ek$)$l$ nesting.} 
\label{tab:sse_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadj$noE$CES$kl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadj$withE$CES$ekl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adj$withE$CES$ekl)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the ($ek$)$l$ nesting, 
the Solow residual is positive, and 
quality-adjusting the inputs 
reduces the Solow residual (\texttt{lambda}).
Adding energy decreases the Solow residual 
(from \Sexpr{naturalCoef(models$PT$unadj$withE$CES$ekl)$lambda} to 
\Sexpr{naturalCoef(models$PT$adj$withE$CES$ekl)$lambda})
when enegy is included in the model.
But, adding energy slightly increases the Solow residual
(from \Sexpr{naturalCoef(models$PT$adj$noE$CES$kl)$lambda} to \Sexpr{naturalCoef(models$PT$adj$withE$CES$ekl)$lambda})
when quality-adjusted inputs are used.

@@

<<UK CES-kl, warning=FALSE>>=
naturalCoef(models$UK$unadj$noE$CES$kl)
naturalCoef(models$UK$adj$noE$CES$kl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals for the UK,
\Sexpr{naturalCoef(models$UK$unadj$noE$CES$kl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$UK$adj$noE$CES$kl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the UK Solow residual.

@@

<<UK CES-kle unadjusted>>=
naturalCoef(models$UK$unadj$withE$CES$kle)
@

@@

<<UK CES-kle quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CES$kle)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CES$kle)$lambda} 
                                                                & \Sexpr{naturalCoef(models$UK$adj$withE$CES$kle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{UK $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadj$noE$CES$kl)$sse} & \Sexpr{EconModels:::sse(models$UK$adj$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadj$withE$CES$kle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adj$withE$CES$kle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<UK CES-lek unadjusted>>=
naturalCoef(models$UK$unadj$withE$CES$lek)
@

@@

<<UK CES-lek quality-adjusted>>=
naturalCoef(models$UK$adj$withE$CES$lek)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadj$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adj$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadj$withE$CES$lek)$lambda} 
                                                            & \Sexpr{naturalCoef(models$UK$adj$withE$CES$lek)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

>>>>

<<setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE>>=
require(EconData)
require(EconModels)
require(dplyr)
require(reshape2)
require(ggplot2)
require(mosaic)

# Tell whether to re-run the analysis to create the models.
# This file will use previously-saved models to generate
# the paper if set to FALSE, greatly reducing time.
# If something changes in the data or in the code that
# performs the analysis, set 
# rerunModels = TRUE
# once to re-generate the information needed to 
# create the paper.
# ********************************************************
rerunModels=FALSE
# ********************************************************

opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
@

@@

historicalPlot(histDataForGraph, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(1,2,4,3,2,4,3,2,3,2,3)) + 
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  coord_cartesian(ylim=c(0.5, 13)) +
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

# Fix levels
pre_econometric_data$flavor <- relevelFactor(as.factor(as.character(pre_econometric_data$flavor)), 
                                               c("Unadjusted", "Quality-adjusted"))
pre_econometric_data$Country <- relevelFactor(as.factor(pre_econometric_data$Country), 
                                              c("PT", "UK"))
pre_econometric_data$variable <- relevelFactor(as.factor(pre_econometric_data$variable),
                                               unique(pre_econometric_data$variable))
# Now make the plot
historicalPlot(pre_econometric_data, 
               mapping=aes(x=Year, y=ratio, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(2,4,3, 2,4,3, 2,3,2, 3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

@@

<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Residuals for CES models with ($le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest == "lek" | nest=="kl")))
@

@@

<<Portugal C-D undadjusted without energy>>=
naturalCoef(models$PT$unadjusted$noE$CD)
@

<<Portugal C-D undadjusted with energy>>=
naturalCoef(models$PT$unadjusted$withE$CD)
@

The Solow residual (\texttt{lambda}) is always negative. 
However, the absolute value of \texttt{lambda} decreases from 
\Sexpr{naturalCoef(models$PT$unadjusted$noE$CD)$lambda} to 
\Sexpr{naturalCoef(models$PT$unadjusted$withE$CD)$lambda}
when adding energy.
So it appears that for Portugal, 
using unadjusted inputs and including energy as a factor of production
reduces the absolute value of the Solow residual somewhat.

@@

<<Portugal C-D adjusted without energy>>=
naturalCoef(models$PT$adjusted$noE$CD)
@

<<Portugal C-D adjusted with energy>>=
naturalCoef(models$PT$adjusted$withE$CD)
@

Using quality-adjusted inputs, the Solow residual (\texttt{lambda}) moves from 
\Sexpr{naturalCoef(models$PT$adjusted$noE$CD)$lambda} to
\Sexpr{naturalCoef(models$PT$adjusted$withE$CD)$lambda} when including energy.
The absolute value of the Solow residual again decreases when adding energy.
But, with quality-adjusted inputs, the effect is much stronger.

@@

\begin{table} \caption{Portugal Solow residual results with Cobb-Douglas.} 
\label{tab:lambda_CD_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadjusted$noE$CD)$lambda}    & \Sexpr{naturalCoef(models$PT$adjusted$noE$CD)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadjusted$withE$CD)$lambda}  & \Sexpr{naturalCoef(models$PT$adjusted$withE$CD)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{UK Solow residual results with Cobb-Douglas.} 
\label{tab:lambda_CD_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadjusted$noE$CD)$lambda}   
                                                & \Sexpr{naturalCoef(models$UK$adjusted$noE$CD)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadjusted$withE$CD)$lambda} 
                                                & \Sexpr{naturalCoef(models$UK$adjusted$withE$CD)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the UK, the quality-adjusted, with energy Solow residual is 
\Sexpr{naturalCoef(models$UK$adjusted$withE$CD)$lambda * 100}\%,
a very small number indeed.
For the UK, it appears that \emph{both} quality-adjustment and addition of energy as a factor of production
are required to effectively eliminate the Solow residual,
reducing it to \Sexpr{naturalCoef(models$UK$adjusted$withE$CD)$lambda}.

Adding energy alone actually \emph{in}creases the Solow residual from 
\Sexpr{naturalCoef(models$UK$unadjusted$noE$CD)$lambda} to
\Sexpr{naturalCoef(models$UK$unadjusted$withE$CD)$lambda}.
The details for each model with unadjusted factors of production are shown below.

<<UK CD Unadjusted Coefficients>>=
naturalCoef(models$UK$unadjusted$noE$CD)
naturalCoef(models$UK$unadjusted$withE$CD)
@

When energy is included, much of the effect of capital (\texttt{alpha\_1}) is 
attributed to energy (\texttt{alpha\_3}), 
as \texttt{alpha\_1} moves from 
\Sexpr{naturalCoef(models$UK$unadjusted$noE$CD)$alpha_1} to
\Sexpr{naturalCoef(models$UK$unadjusted$withE$CD)$alpha_1}.

@@

<<UK CD Quality-adjusted Coefficients>>=
naturalCoef(models$UK$adjusted$noE$CD)
naturalCoef(models$UK$adjusted$withE$CD)
@

\texttt{alpha\_1} moves from 
\Sexpr{naturalCoef(models$UK$adjusted$noE$CD)$alpha_1} to
\Sexpr{naturalCoef(models$UK$adjusted$withE$CD)$alpha_1}.

@@

<<PT CES-kl, warning=FALSE>>=
naturalCoef(models$PT$unadjusted$noE$CES$kl)
naturalCoef(models$PT$adjusted$noE$CES$kl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals,
\Sexpr{naturalCoef(models$PT$unadjusted$noE$CES$kl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the Solow residual.

@@

<<PT CES-kle unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadjusted$withE$CES$kle)
@

@@

<<PT CES-kle quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adjusted$withE$CES$kle)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadjusted$noE$CES$kl)$lambda} 
                                                            & \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$kle)$lambda} 
                                                            & \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$kle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadjusted$noE$CES$kl)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adjusted$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadjusted$withE$CES$kle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$PT$adjusted$withE$CES$kle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<PT CES-lek unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadjusted$withE$CES$lek)
@

@@

<<PT CES-lek quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adjusted$withE$CES$lek)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadjusted$noE$CES$kl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$lek)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$lek)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($le$)$k$ nesting.} 
\label{tab:sse_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadjusted$noE$CES$kl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adjusted$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadjusted$withE$CES$lek)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adjusted$withE$CES$lek)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Similar to the CES model with ($kl$)$e$ nesting,
the CES model with ($le$)$k$ nesting
produces positive Solow residuals in all cases. 
After energy has been added, quality-adjustment 
has very little effect on the Solow residual,
creating a slight increase from 
\Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$lek)$lambda} to
\Sexpr{naturalCoef(models$PT$adjusted$withE$CES$lek)$lambda}.

@@

<<PT CES-ekl unadjusted, warning=FALSE>>=
naturalCoef(models$PT$unadjusted$withE$CES$ekl)
@

@@

<<PT CES-ekl quality-adjusted, warning=FALSE>>=
naturalCoef(models$PT$adjusted$withE$CES$ekl)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$PT$unadjusted$noE$CES$kl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$ekl)$lambda} 
                                        & \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$ekl)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{PT $sse$ with CES and ($ek$)$l$ nesting.} 
\label{tab:sse_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$PT$unadjusted$noE$CES$kl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adjusted$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$PT$unadjusted$withE$CES$ekl)$sse} 
                                        & \Sexpr{EconModels:::sse(models$PT$adjusted$withE$CES$ekl)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For the ($ek$)$l$ nesting, 
the Solow residual is positive, and 
quality-adjusting the inputs 
reduces the Solow residual (\texttt{lambda}).
Adding energy decreases the Solow residual 
(from \Sexpr{naturalCoef(models$PT$unadjusted$withE$CES$ekl)$lambda} to 
\Sexpr{naturalCoef(models$PT$adjusted$withE$CES$ekl)$lambda})
when enegy is included in the model.
But, adding energy slightly increases the Solow residual
(from \Sexpr{naturalCoef(models$PT$adjusted$noE$CES$kl)$lambda} to \Sexpr{naturalCoef(models$PT$adjusted$withE$CES$ekl)$lambda})
when quality-adjusted inputs are used.

@@

<<UK CES-kl, warning=FALSE>>=
naturalCoef(models$UK$unadjusted$noE$CES$kl)
naturalCoef(models$UK$adjusted$noE$CES$kl)
@

Both unadjusted and quality-adjusted variables give positive Solow residuals for the UK,
\Sexpr{naturalCoef(models$UK$unadjusted$noE$CES$kl)$lambda} for unadjusted inputs and
\Sexpr{naturalCoef(models$UK$adjusted$noE$CES$kl)$lambda} for quality-adjusted inputs.
Furthermore, quality-adjustment reduces the magnitude of the UK Solow residual.

@@

<<UK CES-kle unadjusted>>=
naturalCoef(models$UK$unadjusted$withE$CES$kle)
@

@@

<<UK CES-kle quality-adjusted>>=
naturalCoef(models$UK$adjusted$withE$CES$kle)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadjusted$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adjusted$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadjusted$withE$CES$kle)$lambda} 
                                                                & \Sexpr{naturalCoef(models$UK$adjusted$withE$CES$kle)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{UK $sse$ with CES and ($kl$)$e$ nesting.} 
\label{tab:sse_CES-kle_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{EconModels:::sse(models$UK$unadjusted$noE$CES$kl)$sse} & \Sexpr{EconModels:::sse(models$UK$adjusted$noE$CES$kl)$sse} \\
      With energy     & \Sexpr{EconModels:::sse(models$UK$unadjusted$withE$CES$kle)$sse} 
                                                            & \Sexpr{EconModels:::sse(models$UK$adjusted$withE$CES$kle)$sse} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<UK CES-lek unadjusted>>=
naturalCoef(models$UK$unadjusted$withE$CES$lek)
@

@@

<<UK CES-lek quality-adjusted>>=
naturalCoef(models$UK$adjusted$withE$CES$lek)
@

@@

\begin{table} \caption{UK Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_UK} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(models$UK$unadjusted$noE$CES$kl)$lambda} & \Sexpr{naturalCoef(models$UK$adjusted$noE$CES$kl)$lambda} \\
      With energy     & \Sexpr{naturalCoef(models$UK$unadjusted$withE$CES$lek)$lambda} 
                                                            & \Sexpr{naturalCoef(models$UK$adjusted$withE$CES$lek)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Table~\ref{tab:sse_CES-lek_UK} shows the $sse$ values for each of these models.
